Design/Cybernetics: Two Sides of One Coin

### 2019-11-01 MSC-ReadingCourse.tex
final changes

% Refs
\usepackage[backend=bibtex,style=alphabetic]{biblatex}
\addbibresource{refs.bib}

@@

        \begin{frame}
        \frametitle{Prologue II}
        	{\textbf{Gödel's incompleteness} theorems can be summarized as: \\
        		It is impossible to construct a theoretical language that will not \textcolor{teal}{close on itself}, with the result that there are statements that are \textcolor{teal}{undecidable} in that language.} \\
        	\vspace{0.5cm}
        	{Simply put: There are things that you \textcolor{teal}{cannot explain within a paricular language}.}

    \section*{Overview}
    	\begin{frame}
		\frametitle{Agenda}
    		\tableofcontents
    	\end{frame}

    \section{Introduction}

@@

                    \item<1-> Control \neq Restriction
                    \item<2-> Control = Guide towards better performance

@@

                        \item Goal or Intention

@@

                \begin{frame}{Subjective/Objective}
                    \framesubtitle{Observer Incorporated? \cite{vonFoerster_2003}}

@@

                				\includegraphics[height=0.5\textheight]{resources/man.png}

@@

                                    $\longrightarrow x_\infty=f(x_\infty) = {{x_\infty}\over{2}} + 1 = \textcolor{red}{2.000}$\\

@@

                        \item<2-> $2$ is the \textcolor{teal}{Eigen value} of $f$

@@

                        ``This sentence is composed of ... letters'' $\implies$ $\nexists$

@@

@@

@@

                \begin{frame}{Why Conversation? \cite{pangaro_web}}

@@

        \

\section{References}
	\begin{frame}[allowframebreaks]
		\frametitle{References}
		\printbibliography
	\end{frame}

\section*{Appendix}
	\begin{frame}
		\setbeamercovered{transparent}
		\frametitle{Definition(s) of Cybernetics}
		\framesubtitle{Soooo... What is Cybernetics?}
		\begin{itemize}
			\item<1->Ship of the State: The Command of a naval vessel is a metaphor for the governance of a city/state.\\
			--- Plato (420s-340s B.C.)
			\item<2->``Cybernetique is the art of governing or the science of government.''\\
			--- André-Marie Ampère (1775-1836)
			\item<3->``Use the word ‘cybernetics’, Norbert, because nobody knows what it means. This will always put you at an advantage in arguments.''\\
			Widely quoted; attributed to Claude Shannon in a letter to Norbert Wiener in the 1940s.
		\end{itemize}
		\setbeamercovered{invisible}
	\end{frame}
	\begin{frame}{What is NOT a Conversation?}
        \framesubtitle{Shannon's Model of Transmission}
        \begin{figure}
            \centering\includegraphics[width=0.8\textwidth]{resources/transmission.PNG}
        \end{figure}
        This model does not scaffold novelty, is pure mechanical, and does not support learning.
    \end{frame}
    \begin{frame}{Mental Models}
        \framesubtitle{Shaping and re-Shaping}
            \centering\includegraphics<1>[width=0.75\textwidth]{resources/models6.PNG}
            \centering\includegraphics<2>[height=0.77\textheight]{resources/models7.PNG}
    \end{frame}
    \begin{frame}{Interaction}
        \framesubtitle{Configurations \cite{pangaro_web}}
		\begin{figure}
		    \centering
		    \includegraphics[height=0.75\textheight]{resources/interaction.PNG}
		\end{figure}
    \end{frame}
\end{document}

>>>>

% Refs
\usepackage{csquotes}
\usepackage[backend=bibtex,style=alphabetic]{biblatex}
\addbibresource{refs.bib}

@@

        \begin{frame}
        \frametitle{Prologue II}
        	{\textbf{Gödel's incompleteness} theorems can be summarized as: \\
        		It is impossible to construct a theoretical language that will not \textcolor{teal}{close on itself}, with the result that there are statements that are \textcolor{teal}{undecidable} in that language.} \\
        	\vspace{0.5cm}
        	{Simply put: There are things that you \textcolor{teal}{cannot explain within a paricular language}.}
        \end{frame}

    \section{What?}

@@

                    \item<1-> Control $\neq$ Restriction
                    \item<2-> Control $=$ Guide towards better performance\\
                    $\equiv$ Enabling for effectiveness\\
                    $\equiv$ Effective management (Stafford Beer)

@@

                        \item Goal or intention

@@

                    \frametitle{Subjective/Objective}
                    \framesubtitle{Observer Incorporated?}

@@

            				\centering\includegraphics[height=0.5\textheight]{./resources/man.png}\\

@@

                                    $\longrightarrow x_\infty=f(x_\infty) = {{x_\infty}\over{2}} + 1 =$ \textcolor{red}{$2.000$}\\

@@

                        \item<2-> 2 is the \textcolor{teal}{Eigen value} of $f$

@@

                        ``This sentence is composed of ... letters'' $\implies \nexists$

@@

                    \framesubtitle{Viable Systems Model (Stafford Beer)}
                        \column{0.5\textwidth} \centering\includegraphics[height=0.8\textheight]{resources/brain_spinalCord.jpg}
                        \column{0.5\textwidth} \centering\includegraphics[height=0.75\textheight]{resources/VSM.png}

@@

                    \framesubtitle{Abductive Reasoning}

@@

        \subsection{Determinism}
            \begin{frame}{Cybernetics, Design, and Determinism}
            \framesubtitle{Variety and Design: The Undefinable}
                    \item<1-> The vast universe is nevertheless \textcolor{red}{finite}\\
    		        $\implies$ computing capacity \textcolor{red}{always} has a limit (even the most powerful conceivable systems)
    		        \item<2-> Even simple problems can very rapidly become \textcolor{red}{transcomputable}.\\
    		            Ex: configurations of lightbulbs on a 20$\times$20 matrix
    		            Design almost always faces transcomputable problems.
		            \item<3-> The new: something that is not inherent in the existing \\
		            $\implies$ not predictable $\equiv$ not dependent on any variable $\equiv$ not emergent \\
    		        The Undefinable: Impossible to expect to \textcolor{red}{accurately} define a design problem.
            \begin{frame}{Cybernetics, Design, and Determinism}
            \framesubtitle{Variety and Design: Definability and Variety}
    		        \item<1-> Viable Control = Effective Management = Regulation = Stability = Sustainability = ... \\
    		        $\implies$ \textcolor{teal}{Requisite Variety} (controll\textcolor{red}{er} $\leqslant$ controll\textcolor{red}{ed})
    		        \item<2-> Second-Order Cybernetics: Control is emergent (distinctiveness of controlled and controller is arbitrary).\\
        		        $\longrightarrow$ For second-order systems: stability $\equiv$ the variety can only be \textcolor{teal}{the same}.
                    \item<3-> Defining a problem = modeling = controlling (without restricting) the performance of a system\\
                    \item<4-> What if the system (the problem under study) is transcomputable?
                            \item<5-> Changing the definitions of the context under which we chose the states (simplifying)
                            \item<6-> Transforming the notion of control: control-as-restricting
            \begin{frame}{Cybernetics, Design, and Determinism}
            \framesubtitle{Variety and Design: Unmanageability and Creativity}
    		        \item<1-> Being in control = Defining the range of possible (what will be considered)
    		        \item<2-> When I am in control $\implies$ I restrict the world to what I can imagine or permit\\
    		            $\implies$ I support a predetermination of what-is and what-might-be and aim towards predetermined goals.
		            \item<3-> Ex: Deciding to go to a restaurant:\\
		                    If I choose (restrict) $\longrightarrow$ reflections of my taste, knowledge and ignorance.\\
		                    If I let others choose $\longrightarrow$ I can find new experiences.
                    \item<4-> Benefit of not having enough variety to control the system\\
                    $\longrightarrow$ Discover \textcolor{teal}{the unexpected} and \textcolor{teal}{the novel}
                    \item<5-> Not restricting the future considerations to current knowledge: Experiencing the vast unknown
        \subsection{Design As Done}
            \begin{frame}{Design As Done}
            \framesubtitle{Design as a Conversation with self (and others)}
    		        \item<1-> We look and then we draw
    		        \item<1-> We see something new, not previously intended
    		        \item<2-> Conversation requires switching roles: \textcolor{teal}{pendulum}
    		        \item<3-> Meanings are private $\implies$ understandings are different: \textcolor{teal}{novelty}
    		        \item<4-> Agree to agree; or agree to disagree
    		        \item<4-> Test for understanding: Consistency of what I hear and what I say
    		        \item<5-> $\longrightarrow$ Increasing the \textcolor{teal}{variety} of the ``repertoire'' of the designer
    		        \item<6-> Recursion, role switching $\longrightarrow$ understand \textcolor{red}{well enough} to face \textcolor{red}{complexity}
		            \item<7-> We \textcolor{teal}{construct} our understanding
            \begin{frame}{Design As Done}
            \framesubtitle{Conversation, Objects, Eigen forms, and Autonomy}
    		        \item<1-> Conversation with self
    		        \item<2-> Autopoietic systems: Create conditions to create themselves; continue to \textcolor{teal}{reproduce} themselves in the environment
    		        \item<3-> Eigen forms: a system can operate on its own output, \textcolor{teal}{treating} it as an input.
    		        \item<4-> Designers: create the conditions in which the design outcome can come into being; continue the design act to reach a constant outcome (details  may be enriched).
    		        \item<5-> Children: grow to become their own persons
    		        \item<6-> $\longrightarrow$ \textcolor{red}{Autonomy = Organizational Closure}
            \begin{frame}{Design As Done}
            \framesubtitle{Conversation as Design}
    		        \item<1-> Design translated once more: \textcolor{red}{Wandering}
                    \item<2-> Once arrived, we can make sense of the progress. (the ``solution'' defines the ``problem'')
                    \item<3-> Arriving and stopping? Eigen forms? \textcolor{teal}{Well enough: stable form}.
                    \item<4-> Stopping and re-starting? 
                    \item<5-> Pendulum: For a second-order description, observers \textcolor{red}{out} must become observers \textcolor{red}{in}.
            \begin{frame}{Design As Done}
            \framesubtitle{Being Out of Control}
    		        \item<1-> Not being in control $\implies$ a way of increasing creativity
                    \item<2-> Restricting responses and conversation to what I know\\
                        $\implies$ Destroying conversation
                        $\implies$ Conversation controlled by a participant $\longrightarrow \nexists$
                    \item<3-> Design is a second-order activity: Ashby's law cannot be utilised.
                    \item<4-> Stopping and re-starting? 
                    \item<5-> Pendulum: For a second-order description, observers \textcolor{red}{out} must become observers \textcolor{red}{in}.
            \begin{frame}{Design As Done}
            \framesubtitle{Complexity}
    		        \item<1-> What if the system (the problem under study) is transcomputable?
                            \item Changing the definitions of the context (simplifying)
                            \item Transforming the notion of control: control-as-restricting
    		        \item<1-> Designer: someone who faces \textcolor{red}{complexity} (and ambiguity)
                    \item<2-> ``The systems theorist of the future, I suggest, must be an expert in how to simplify'' \\--- Ross Ashby
                    \item<3-> ``Progress means simplifying, not complicating'' --- Bruno Munari (Italian Designer)
                    \item<4-> Simplifying: a process by which complex requirements can be brought together within \textcolor{teal}{one}, \textcolor{teal}{unified}, \textcolor{teal}{unitary} form.
                    \item<4-> Design brings to complexity an approach that is distinct from complexity science, which can lead to added value.
        \subsection{Criteria and Conditions}
            \begin{frame}{From Cybernetics to Design}
    		        \item<1-> Design leads not to the best, but to a large variety of different outcomes, giving \textcolor{red}{choice}.
    		        \item<2-> The nature of design conversation:\\
    		        For it to operate $\implies \exists$ a listener (viewer)\\
    		        To listen (see) $\implies \exists$ an open (and receptive) mind, and generosity\\
    		        To design $\implies$ see the possibilities \textcolor{red}{not already existent} in the mind
    		        To increase choices through design $\implies$ acting responsibly
    		        \item<3-> Cybernetic systems work if we accept responsibility, and act generously, with an open mind.
    		        \item<4-> Conversation is the intersection
        \subsection{Notes on Epistemology}
            \begin{frame}{Knowledge Of and Knowledge For}
    		        \item<1-> Science \textcolor{red}{through an observer} gives us \textcolor{red}{knowledge of} the system. \\
    		        Helps us understand \textcolor{red}{what is} $\longrightarrow$ passive and neutral \\
    		        $\implies$ \textcolor{orange}{Tacit} knowledge
    		        \item<2-> Designers are actors \textcolor{red}{in} the world $\longrightarrow$ require and generate \textcolor{red}{knowledge for} acting in the world as they make it.\\
    		        $\implies$ \textcolor{teal}{Reflective} knowledge
    		        \item<3-> Technology (engineering): converts `knowledge of' into `knowledge for'
    \section{Discussion}
        \begin{frame}{Wicked Problems}
		    \centering\includegraphics[height=0.8\textheight]{resources/rittle.png}
        \begin{frame}{Cybernetics and Design: Conversations for Action \cite{Dubberly2019}}
		    \centering\includegraphics<1>[page=31, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<2>[page=32, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<3>[page=33, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<4>[page=34, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<5>[page=35, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<6>[page=36, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<7>[page=37, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<8>[page=38, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
		    \centering\includegraphics<9>[page=39, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
        \begin{frame}{Interaction \cite{pangaro_web}}
    		    \includegraphics<1>[height=0.75\textheight]{resources/interaction.PNG}
    		    \includegraphics<2>[height=0.75\textheight]{resources/conversing1.png}
    		    \includegraphics<3>[height=0.75\textheight]{resources/conversing2.png}
        \begin{frame}{Conversations? \cite{pangaro_web}}

@@

        \begin{frame}{\cite{pangaro_web}}
            \centering\includegraphics<1>[page=41, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<2>[page=42, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<3>[page=44, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<4>[page=89, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<5>[page=90, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
        \begin{frame}{Implications \cite{pangaro_web}}
            \centering\includegraphics<1>[page=92, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<2>[page=93, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<3>[page=94, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<4>[page=95, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<5>[page=96, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<6>[page=97, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<7>[page=98, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<8>[page=99, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<9>[page=100, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<10>[page=101, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<11>[page=102, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<12>[page=103, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<13>[page=104, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<14>[page=105, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<15>[page=106, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<16>[page=107, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<17>[page=108, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<18>[page=109, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<19>[page=110, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<20>[page=111, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<21>[page=112, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<22>[page=113, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<23>[page=114, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<24>[page=115, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<25>[page=127, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
            \centering\includegraphics<26>[page=128, width=0.75\textwidth]{resources/Pangaro_Vienna-HvF-Lecture_June-2017_distro-r.pdf}
        \begin{frame}{Design of Ethical Interfaces \cite{pangaro_web}}
            \centering\includegraphics<1>[page=12, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<2>[page=13, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<3>[page=14, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<4>[page=15, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<5>[page=16, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<6>[page=17, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<7>[page=49, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<8>[page=18, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<9>[page=19, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<10>[page=20, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}
            \centering\includegraphics<11>[page=21, width=0.75\textwidth]{resources/Pangaro-HCII_Seminar-April_2019-distro.pdf}

    \section*{Overview}
    	\begin{frame}
		\frametitle{Overview}
    		\tableofcontents
    	\end{frame}

\section*{References}
	\begin{frame}[allowframebreaks]
		\frametitle{References}
		\printbibliography
	\end{frame}

\section*{Appendix}
	\begin{frame}
		\setbeamercovered{transparent}
		\frametitle{Definition(s) of Cybernetics}
		\framesubtitle{Soooo... What is Cybernetics?}
		\begin{itemize}
			\item<1->Ship of the State: The Command of a naval vessel is a metaphor for the governance of a city/state.\\
			--- Plato (420s-340s B.C.)
			\item<2->``Cybernetique is the art of governing or the science of government.''\\
			--- André-Marie Ampère (1775-1836)
			\item<3->``Use the word ‘cybernetics’, Norbert, because nobody knows what it means. This will always put you at an advantage in arguments.''\\
			Widely quoted; attributed to Claude Shannon in a letter to Norbert Wiener in the 1940s.
		\end{itemize}
		\setbeamercovered{invisible}
	\end{frame}
	\begin{frame}{What is NOT a Conversation?}
        \framesubtitle{Shannon's Model of Transmission}
        \begin{figure}
            \centering\includegraphics[width=0.8\textwidth]{resources/transmission.PNG}
        \end{figure}
        This model does not scaffold novelty, is pure mechanical, and does not support learning.
    \end{frame}
    \begin{frame}{Interaction}
        \framesubtitle{Forming an interaction - Meaning is PRIVATE}
        \centering\includegraphics[height=0.75\textheight]{resources/interaction1.png}
    \end{frame}
    \begin{frame}{Relationship}
        \framesubtitle{Interactions Flowing in Both Direction + Time + Recursion = TRUST}
        \centering\includegraphics[height=0.75\textheight]{resources/interaction2.png}
    \end{frame}
    \begin{frame}{Conversation}
        \framesubtitle{``Immaterial'' Aspects}
        \centering\includegraphics[height=0.75\textheight]{resources/interaction3.png}
    \end{frame}
    \begin{frame}{Mental Models}
        \framesubtitle{Shaping and re-Shaping}
            \centering\includegraphics<1>[width=0.75\textwidth]{resources/models6.PNG}
            \centering\includegraphics<2>[height=0.77\textheight]{resources/models7.PNG}
    \end{frame}
\end{document}
